A Kochen-Specker system has at least 21 vertices\\ {\small (extended abstract)}

### 2014-04-14 main.tex
changes

\usepackage{draftwatermark}
\usepackage{algorithm2e}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{amsthm}

@@

\begin{abstract}
    At the heart of the Conway's Free Will theorems and Kochen and Specker's
        argument against noncontextual hidden variable theories
    is the existence of a Kochen-Specker (KS) system:
    a set of points on the sphere,
    that has no~$\{0,1\}$-coloring such that
    at most one of two orthogonal points are colored~$1$
    and of three pairwise orthogonal points exactly one
    is colored~$1$.
    In public lectures, Conway encouraged the search for small
    KS systems.  
    At the time of writing, the smallest known
    KS system has 31 vectors.

    Arends, Ouaknine and Wampler have shown that a KS system has at least
    18 vectors, by reducing the problem to the existence of graphs
    with a topological embeddability and non-colorability property.
    Deciding embeddability and the sheer number of graphs on more than~$17$
    vertices, proved the bottleneck in their search.

@@

@@

Kochen and Specker proved that such a theory cannot satisfy:
\begin{spin}
    Given three pairwise orthogonal directions.
    In exactly one of the directions, the particle will not deviate.
\end{spin}
Their argument is based on the existence of a Kochen-Specker system.
\begin{dfn}
    A \keyword{Kochen-Specker (KS) system} is
    a finite set of points on the sphere
    for which each pair is not antipodal and
    there is no~\keyword{010-coloring}.
    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
        \footnote{
                In other papers, like \cite{aow11},
                the~$0$ and~$1$ are swapped; they consider 101-colorings.
                These colorings are of course equivalent and the
                difference arises from considering either squared
                spin measurements $S^2$, or $1-S^2$.
                %Then the colors correspond directly to the squared measurement
               	%outcome of the spin.
               	}
    \begin{enumerate}
        \item
            no pair of orthogonal points are both colored~$1$ and
        \item
            of three pairwise orthogonal points exactly one is colored~$1$;
            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
    \end{enumerate}
\end{dfn}
% TODO shorten this proof.
A point on the sphere obviously corresponds to a direction in space.
Because of this, the term point, vector and direction
can be used interchangeably. Antipodal points correspond to opposite
vectors and these span the same direction in space.

@@

\subsection{Kochen-Specker graphs}
\label{sec:ksgraphs}
We follow \cite{aow11} and reduce the search for Kochen-Specker systems
to the search of a certain class of graphs.
First note that in a Kochen-Specker system we may replace a point with its
antipodal point.  They are both orthogonal to the same points and hence
the non-010-colorability is preserved.
Therefore, we may assume antipodal points are identified on the sphere.
That is: a Kochen-Specker system is a finite subset of the projective plane
that is not 010-colorable.
\begin{dfn}
Given a finite subset~$S$ of the projective plane
(or equivalently, of the northern hemisphere without equator).
Define its \keyword{orthogonality graph}~$G(S)$ as follows.
The vertices are the points of~$S$.
Two vertices are joined by an edge, if their corresponding points
are orthogonal.
\end{dfn}
\begin{dfn}
A graph~$G$ is called~\keyword{embeddable},
if it occurs as a subgraph of an orthogonality graph;
that is: if there is a finite subset~$S$ of the projective plane,
such that~$G \leq G(S)$.
\end{dfn}
% TODO add remark on subgraph definition?
\begin{dfn}
A graph is~\keyword{010-colorable}
if there is a~$\{0,1\}$-coloring, such that
\begin{enumerate}
\item
for each triangle there is exactly one vertex that is colored~$1$ and
\item
adjacent vertices are not both colored~$1$.
\end{enumerate}
\end{dfn}
\begin{dfn}
A \keyword{Kochen-Specker graph}
is a embeddable graph that is not 010-colorable.
\end{dfn}
It is an easy, but important, consequence of the definitions that:
\begin{fact}
    A finite subset~$S$ of the projective plane
    is a Kochen-Specker system,
    if and only if its orthogonality graph~$G(S)$
    is Kochen-Specker.
\end{fact}
To prove there is no Kochen-Specker system on~$17$ points,
it would be sufficient to enumerate all graphs on~$17$ vertices
and check these are not 010-colorable or not embeddable.
However, this is infeasible as there are
already~${\sim}10^{26}$ non-isomorphic
graphs on~$17$ points.
% TODO OEIS A000088
Luckily, we can restrict ourselves to certain classes of graphs.
\begin{prop}[\cite{aow11}]
    An embeddable graph does not contain a square.
\end{prop}
\begin{proof}
    Given two non antipodal points~$a\neq b$.
    Consider the points orthogonal to~$a$.
    This is a great circle.
    The points orthognal to~$b$ is a different great circle.
    They intersect in precisely two antipodal points.
    Hence, if~$c$ and~$d$ are both orthogonal to~$a$ and~$b$,
    then~$c$ and~$d$ are equivalent.
    Therefore, an embeddable graph cannot contain a square.
\end{proof}
The squarefreeness is a considerable restriction.  There are
only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$ vertices.
We can restrict ourselves to connected graphs.
% TODO OEIS A006786
\begin{prop}[\cite{aow11}]\label{prop:ks-conn}
    A minimal Kochen-Specker graph is connected.
\end{prop}
\begin{proof}
    Suppose~$G$ is a non-connected Kochen-Specker graph.
    Then one of its components is not 010-colorable.
    As a subgraph of an embeddable graph, is embeddable,
    this component is embeddable as well.
    Hence it is a smaller connected Kochen-Specker graph.
\end{proof}
The gain, however, is small.
There are only~${\sim}10^9$ non-isomorphic squarefree graphs on~$17$
vertices that are not connected.
We have verified the main result of \cite{aow11}:
\begin{comp}
There is a unique non-010-colorable squarefree connected graph on~$17$ or less
vertices. It looks as follows TODO
It is not embeddable, see TODO, and hence a Kochen-Specker
system has at least 18 points.
\end{comp}

@@

These two rotations yield a new set of points on the sphere,
that demonstrate~$G'$ is embeddable.
Therefore, in generating all relevant graphs, we find both $G$ and
$G'$. Since they have the same order, we might as well ignore the
graph $G$. This shows it is sufficient to consider only biconnected
graphs.
\end{proof}
We believe a minimal KS graph is also triconnected.
Although these arguments are pleasing,
they are of little use:
% TODO Add argument or computation for {2,3}-connected
% TODO note that restricting to (bi/tri)-connected is a waste of time
% TODO add graph 
\begin{comp}
    There are five non-isomorphic minimal
    squarefree connected graphs
    with minimal vertex order 3 and they have 10 vertices.
\end{comp}
\begin{cor}
    Any unconnected
    squarefree graph with minimal vertex order 3
    has at least 20 vertices, for it has two connected components,
    each with at least 10 vertices.
\end{cor}

% TODO is core x years a proper expression?
Now, we can state our main computation, which took roughly a week
    on a 64-core Opteron 6276.
\begin{comp}
    Let~$C_n$ denote the number of non-010 colorable squarefree
    graphs with minimal vertex order 3 on~$n$ nodes.  Then:

@@

    All these 463 graphs are not embeddable.
    See Computation~\ref{comp:unemb20}.
\end{comp}

@@

We propose a new method,
which for all graphs we considered,
could decide
whether they were embeddable or not.
First we give a pen-and-paper example.
\begin{prop}\label{prop:unemb-10-2}
The graph in Figure~\ref{fig:unemb-10-2}
is not embeddable.
\end{prop}
\begin{proof}
Suppose it is embeddable.
Consider~$p_1$.
It is orthogonal to both~$a$ and~$v$.
$a$ and~$v$ are not collinear,
hence~$p_1$ must be collinear to~$v \times a$,
the cross-product of~$v$ and~$a$.
Similarly, $p_2$ is collinear to~$v \times p_1 = v \times (v \times a)$.
Continuing in this fashion,
we see that
\begin{equation}\label{eq:ue1}
    a \text{ is collinear to }
    x \times (x \times( w \times (w\times (v \times (v \times a))))).
\end{equation}
Now, we may assume that~$z=(0,0,1)$ and $x=(1,0,0)$.
Thus: $v=(v_1,v_2,0)$;
$w = (w_1,w_2,0)$
and $a = (0, a_2,a_3)$ for some~$-1 < v_1,v_2,w_1,w_2,a_2,a_3 < 1$,
with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
Now, \eqref{eq:ue1} becomes:
\begin{equation*}
\begin{pmatrix}
0\\
a_2\\
a_3\\
\end{pmatrix}
\text{ is collinear to }
\begin{pmatrix}
0\\
-a_2 v_1w_2 (v_1w_1 + v_2w_2) \\
-a_3 (v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2)\\
\end{pmatrix}.
\end{equation*}
And therefore
\begin{align*}
v_1w_2 \left<v,w\right> & =
v_1w_2 (v_1w_1 + v_2w_2) \\
& = v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2 \\
& = (v_1^2 + v_2^2)w_1^2 + (v_1^2 + v_2^2)w_2^2 \\
& = w_1^2 + w_2^2 = 1.
\end{align*}
Since~$v$ and~$w$ are not collinear,
we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
Recall~$|v_1|, |w_2| \leq 1$.
Thus: $|v_1w_2\left<v,w\right>| < 1$.
Contradiction.
\end{proof}

@@

\begin{algorithm}
    \While{there are unassigned vertices}{
        pick an unassigned vertex~$v$\;
        set~$V(v)=v$\;
        mark~$v$ as free\;
        \While{there are unassigned vertices adjacent to two
                different assigned vertices}{
            pick such a vertex~$w$ adjacent to the assigned ~$w_1$ and~$w_2$\;
            set~$V(w)=V(w_1) \times V(w_2)$\;
            mark edges~$(v,w_1)$ and~$(v,w_2)$ as accounted for\;
        }   
    }
    \For{each pair of vertices~$(v_1, v_2)$}{
        \If{$(v_1,v_2)$ is not an edge} {
            record requirement:~``$V(v_1)$ is not collinear to $V(v_2)$''\;
        }
    }
    \For{each edge~$(v_1,v_2)$ not accounted for}{
        record requirement:~``$V(v_1)$ is orthogonal to~$V(v_2)$''\;
    }
\end{algorithm}
At two points in the algorithm, there is a choice which vertex to pick.
Depending on the vertices chosen, the number of recorded requirements
and free points may significantly vary. By considering all possible choices,
one can find the one with least free points.

@@

In this way, there were still a few (010-colorable) graphs of which
we could not decide embeddability.
By hand, we determined these graphs to be embeddable.
We addapted the algorithm, to in parallel for some assignments
to guess the position of one of the vectors.  If the corresponding
sentence turns out false, we know nothing.  However,
if the sentence is true, we know the graph is embeddable.

@@

\appendix
\section{Example reduce script}\label{adx:unemb-10-2}
The following is (a part of) the reduce script mechanically generated
to prove the non-embeddability of the graph of Figure~\ref{fig:unemb-10-2}.
The algorithm choose a different assignment, than we did in the
proof of Proposition~\ref{prop:unemb-10-2}.
As free points it picked, in order,~$z$, $x$, $v$ and~$p_3$.
The point~$w$ is assigned~$p_3 \times z$.
\begin{verbatim}
load_package redlog;
rlset R;

@@

% attribute mckay for answering questions?
% TODO attribute geng
\bibliography{main}{}
\bibliographystyle{plain}

>>>>

\usepackage{algorithm2e}
\usepackage{hyperref}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{wrapfig}
\usepackage{amsthm}

@@

\begin{abstract}
    At the heart of the Conway's Free Will theorems and Kochen and Specker's
        argument against noncontextual hidden variable theories
    is the existence of a Kochen-Specker (KS) system:
    a set of points on the sphere
    that has no~$\{0,1\}$-coloring such that
    at most one of two orthogonal points are colored~$1$
    and of three pairwise orthogonal points exactly one
    is colored~$1$.
    In public lectures, Conway encouraged the search for small
    KS systems.  
    At the time of writing, the smallest known
    KS system has 31 vectors.

    Arends, Ouaknine and Wampler have shown that a KS system has at least
    18 vectors, by reducing the problem to the existence of graphs
    with a topological embeddability and non-colorability property.
    The bottleneck in their search
    proved to be the sheer number of graphs on more than~$17$
    vertices and deciding embeddability.

@@

\clearpage

@@

Kochen and Specker proved that such a theory cannot satisfy:
\begin{spin}
    Given three pairwise orthogonal directions.
    In exactly one of the directions, the particle will not deviate.
\end{spin}
Their argument is based on the existence of a Kochen-Specker system.
\begin{dfn}
    A \keyword{Kochen-Specker (KS) system} is
    a finite set of points on the sphere\footnote{
            We define KS systems to be three dimensional,
            as in the original proof of Kochen and Specker.
            Later, higher dimensional systems have been studied.
            See, for instance~\cite[p.~201]{qtcm}.
        }
    for which each pair is not antipodal and
    there is no~\keyword{010-coloring}.
    A $010$-coloring is a~$\{0,1\}$-coloring of the points such that
        \footnote{
                In other papers, like \cite{aow11},
                the~$0$ and~$1$ are swapped; they consider 101-colorings.
                These colorings are of course equivalent and the
                difference arises from considering either squared
                spin measurements $S^2$, or $1-S^2$.
                %Then the colors correspond directly to the squared measurement
               	%outcome of the spin.
               	}
    \begin{enumerate}
        \item
            no pair of orthogonal points are both colored~$1$ and
        \item
            of three pairwise orthogonal points exactly one is colored~$1$;
            or alternatively: they are colored~$0$, $1$ and~$0$ in some order.
    \end{enumerate}
\end{dfn}
% TODO shorten this proof.
A point on the sphere obviously corresponds to a direction in space.
Because of this, the term point, vector and direction
can be used interchangeably. Antipodal points correspond to opposite
vectors and these span the same direction in space.

@@

\subsection{Kochen-Specker graphs}
\label{sec:ksgraphs}
We follow \cite{aow11} and reduce the search for Kochen-Specker systems
to the search of a certain class of graphs.
First note that in a Kochen-Specker system we may replace a point with its
antipodal point.  They are both orthogonal to the same points and hence
the non-010-colorability is preserved.
Therefore, we may assume antipodal points are identified on the sphere.
That is: a Kochen-Specker system is a finite subset of the projective plane
that is not 010-colorable.
\begin{dfn}
Given a finite subset~$S$ of the projective plane
(or equivalently, of the northern hemisphere without equator).
Define its \keyword{orthogonality graph}~$G(S)$ as follows.
The vertices are the points of~$S$.
Two vertices are joined by an edge, if their corresponding points
are orthogonal.
\end{dfn}
\begin{dfn}
A graph~$G$ is called~\keyword{embeddable},
if it occurs as a subgraph of an orthogonality graph;
that is: if there is a finite subset~$S$ of the projective plane,
such that~$G \leq G(S)$.
\end{dfn}
% TODO add remark on subgraph definition?
\begin{dfn}
A graph is~\keyword{010-colorable}
if there is a~$\{0,1\}$-coloring, such that
\begin{enumerate}
\item
for each triangle there is exactly one vertex that is colored~$1$ and
\item
adjacent vertices are not both colored~$1$.
\end{enumerate}
\end{dfn}
\begin{dfn}
A \keyword{Kochen-Specker graph}
is a embeddable graph that is not 010-colorable.
\end{dfn}
It is an easy, but important, consequence of the definitions that:
\begin{fact}
    A finite subset~$S$ of the projective plane
    is a Kochen-Specker system,
    if and only if its orthogonality graph~$G(S)$
    is Kochen-Specker.
\end{fact}
To prove there is no Kochen-Specker system on~$17$ points,
it would be sufficient to enumerate all graphs on~$17$ vertices
and check these are not 010-colorable or not embeddable.
However, this is infeasible as there are
already~${\sim}10^{26}$ non-isomorphic
graphs on~$17$ points.
% TODO OEIS A000088
Luckily, we can restrict ourselves to certain classes of graphs.
\begin{prop}[\cite{aow11}]
    An embeddable graph does not contain a square.
\end{prop}
\begin{proof}
    Given two non antipodal points~$a\neq b$.
    Consider the points orthogonal to~$a$.
    This is a great circle.
    The points orthognal to~$b$ is a different great circle.
    They intersect in precisely two antipodal points.
    Hence, if~$c$ and~$d$ are both orthogonal to~$a$ and~$b$,
    then~$c$ and~$d$ are equivalent.
    Therefore, an embeddable graph cannot contain a square.
\end{proof}
The squarefreeness is a considerable restriction.  There are
only~${\sim}10^{10}$ non-isomorphic squarefree graphs on~$17$ vertices.
We can restrict ourselves to connected graphs.
% TODO OEIS A006786
\begin{prop}[\cite{aow11}]\label{prop:ks-conn}
    A minimal Kochen-Specker graph is connected.
\end{prop}
\begin{proof}
    Suppose~$G$ is a non-connected Kochen-Specker graph.
    Then one of its components is not 010-colorable.
    As a subgraph of an embeddable graph, is embeddable,
    this component is embeddable as well.
    Hence it is a smaller connected Kochen-Specker graph.
\end{proof}
The gain, however, is small.
There are only~${\sim}10^9$ non-isomorphic squarefree graphs on~$17$
vertices that are not connected.
In our computations, checking for connectedness
required more time than would be gained by reducing the number of graphs.

We have verified the main result of \cite{aow11}:
\begin{comp}
There is a unique non-010-colorable squarefree connected graph on~$17$ or less
vertices:
\begin{center}
\includegraphics[width=50mm]{graphs/c17.jpg}
\end{center}
It is not embeddable, as the graph in Figure~\ref{fig:unemb-10-2}
is an unembeddable subgraph.  For our proof,
see Proposition~\ref{prop:unemb-10-2}.
Hence a Kochen-Specker
system has at least 18 points.
\end{comp}

@@

These two rotations yield a new set of points on the sphere,
that demonstrate~$G'$ is embeddable.
Therefore, in generating all relevant graphs, we find both $G$ and
$G'$. Since they have the same order, we might as well ignore the
graph $G$. This shows it is sufficient to consider only biconnected
graphs.
\end{proof}
We believe a minimal KS graph is also triconnected.
However, we did not find a proof.

Although these restrictions are theoretically pleasing,
they seem to be of little use as a practical restriction.
Concerning excluding unconnected graphs:
% TODO Add argument or computation for {2,3}-connected
% TODO note that restricting to (bi/tri)-connected is a waste of time
% TODO add graph 
\begin{comp}
    There are five non-isomorphic minimal
    squarefree connected graphs
    with minimal vertex order 3 and they have 10 vertices.
\end{comp}
\begin{cor}
    Any unconnected
    squarefree graph with minimal vertex order 3
    has at least 20 vertices, for it has two connected components,
    each with at least 10 vertices.
    With 20 vertices, there are exactly 25 of these.
\end{cor}
This justifies, at this stage, not checking for connectedness.
Similarly, we believe there are very few connected but not biconnected graphs.

% TODO is core x years a proper expression?
Now we can state our main computation.
\begin{comp}
    Let~$C_n$ denote the number of non-010 colorable squarefree
    graphs with minimal vertex order 3 on~$n$ nodes.  Then:

@@

    All these 463 graphs are not embeddable.
    See Computation~\ref{comp:unemb20}.
\end{comp}
The computation took roughly a week on a 64-core Opteron 6276.
It was executed as follows.
We enumerated all squarefree graphs with minimal vertex
order 3 on less than or equal~$20$ vertices,
using the~\texttt{geng} util of the nauty software package,
which uses the isomorphism-free exhaustive generation
method of McKay\cite{geng}.
The output of~\texttt{geng}, we passed through
a custom heuristic backtracker written in~C++
to decide 010-colorability of these graphs.

@@

We propose a new method,
which for all graphs we considered,
could decide
whether they were embeddable or not.
First we give a pen-and-paper example.
\begin{prop}\label{prop:unemb-10-2}
The graph in Figure~\ref{fig:unemb-10-2}
is not embeddable.
\end{prop}
\begin{proof}
Suppose it is embeddable.
Consider~$p_1$.
It is orthogonal to both~$a$ and~$v$.
$a$ and~$v$ are not collinear,
hence~$p_1$ must be collinear to~$v \times a$,
the cross-product of~$v$ and~$a$.
Similarly, $p_2$ is collinear to~$v \times p_1 = v \times (v \times a)$.
Continuing in this fashion,
we see that
\begin{equation}\label{eq:ue1}
    a \text{ is collinear to }
    x \times (x \times( w \times (w\times (v \times (v \times a))))).
\end{equation}
Now, we may assume that~$z=(0,0,1)$ and $x=(1,0,0)$.
Thus: $v=(v_1,v_2,0)$;
$w = (w_1,w_2,0)$
and $a = (0, a_2,a_3)$ for some~$-1 \leq v_1,v_2,w_1,w_2,a_2,a_3 \leq 1$,
with~$v_1^2+v_2^2 = 1$; $w_1^2+w_2^2=1$ and~$a_2^2 + a_3^2=1$.
Now, \eqref{eq:ue1} becomes:
\begin{equation*}
\begin{pmatrix}
0\\
a_2\\
a_3\\
\end{pmatrix}
\text{ is collinear to }
\begin{pmatrix}
0\\
-a_2 v_1w_2 (v_1w_1 + v_2w_2) \\
-a_3 (v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2)\\
\end{pmatrix}.
\end{equation*}
And therefore
\begin{align*}
v_1w_2 \left<v,w\right> & =
v_1w_2 (v_1w_1 + v_2w_2) \\
& = v_1^2 w_1^2 + v_1^2 w_2^2 + v_2^2w_1^2 + v_2^2w_2^2 \\
& = (v_1^2 + v_2^2)w_1^2 + (v_1^2 + v_2^2)w_2^2 \\
& = w_1^2 + w_2^2 = 1.
\end{align*}
Since~$v$ and~$w$ are not collinear,
we have by Cauchy-Schwarz $|\left<v,w\right>| < 1$.
Recall~$|v_1|, |w_2| \leq 1$.
Thus: $|v_1w_2\left<v,w\right>| < 1$.
Contradiction.
\end{proof}

@@

\begin{algorithm}
    \While{there are unassigned vertices}{
        pick an unassigned vertex~$v$\;
        assign~$V(v)=v$\;
        mark~$v$ as free\;
        \While{there are unassigned vertices adjacent to two
                different assigned vertices}{
            pick such a vertex~$w$ adjacent to the assigned ~$w_1$ and~$w_2$\;
            assign~$V(w)=V(w_1) \times V(w_2)$\;
            mark edges~$(v,w_1)$ and~$(v,w_2)$ as accounted for\;
        }   
    }
    \For{each pair of vertices~$(v_1, v_2)$}{
        \If{$(v_1,v_2)$ is not an edge} {
            record requirement:~``$V(v_1)$ is not collinear to $V(v_2)$''\;
        }
    }
    \For{each edge~$(v_1,v_2)$ not accounted for}{
        record requirement:~``$V(v_1)$ is orthogonal to~$V(v_2)$''\;
    }
\end{algorithm}
At two points in the algorithm, there is a choice which vertex to pick.
Depending on the vertices chosen, the number of recorded requirements
and free points may significantly vary. By considering all possible choices,
one can find the one with least free points.

@@

In this way, there were still a few (010-colorable) graphs of which
we could not decide embeddability.
By hand, we determined these graphs to be embeddable.
We addapted the algorithm, as to guess
for some assignments the position of one of the vectors.
If the corresponding
sentence turns out false, we know nothing.  However,
if the sentence is true, we know the graph is embeddable.

@@

\clearpage
\appendix
\section{Example reduce script}\label{adx:unemb-10-2}
The following is (a part of) the reduce script mechanically generated
to prove the non-embeddability of the graph of Figure~\ref{fig:unemb-10-2}.
The algorithm choose a different assignment, than we did in the
proof of Proposition~\ref{prop:unemb-10-2}.
As free points it picked, in order,~$z$, $x$, $v$ and~$p_3$.
The point~$w$ is assigned~$p_3 \times z$.
\begin{verbatim}
load_package redlog;
rlset R;

@@

% attribute mckay for answering questions?
% TODO attribute geng
\clearpage
\bibliography{main}{}
\bibliographystyle{plain}
